---
title: "data_manipulation"
---

```{r load-packages}
#| label: load-pkgs
#| message: false
#| warning: false

library(tidyverse)
library(dplyr)
library(tidytuesdayR)
library(lubridate)
library(readr)

# read data 
static_list <-readRDS("data/static_list.RDS")
contribution_data_all_states <-readRDS("data/contribution_data_all_states.RDS")
```


```{r question-1-data-processing}
static_list <- static_list |> 
  mutate(amount_contributed = `Amount Contributed Across States`)

static_list <- static_list |> 
  mutate(hrc_pledge = `HRC Business Pledge`)

static_list <- static_list |> 
  mutate(pride_or_not = `Pride?`)


new_static_list <- static_list |> 
  filter(!is.na(Company), !is.na(hrc_pledge)) |> 
  select(Company, amount_contributed, hrc_pledge, pride_or_not) |> 
  group_by(Company) |> 
  summarise(total_amount = sum(amount_contributed),
            hrc_pledge, pride_or_not) |> 
  arrange(desc(total_amount)) |> 
  slice_head(n = 20)
  
new_static_list |> 
  filter(hrc_pledge == TRUE, pride_or_not == FALSE)
```

For our data processing, we created a year variable by extracting the year from the Date variable. We also filtered for companies and dates that weren't NA; renamed the Pride? variable to pride_or_not for better readability; and looked at the total contribution for both pride and anti-LGBTQ+ campaigns over the years 2013-2022. It is notable that there was no data on contributions towards anti-LGBTQ+ campaigns in 2016.

```{r question-2-data-processing}
company_contribution_all_q2 <- contribution_data_all_states %>%
  filter(!is.na(Company), !is.na(Date)) %>%
  mutate(year = year(ymd(Date)),
         pride_or_not = case_when(`Pride?` == TRUE ~ "Pride",
                                TRUE ~ "Anti-LGBTQ+")) %>%
  group_by(pride_or_not, year) %>%
  summarise(total_contribution = sum(Amount), .groups = 'drop')

company_contribution_all_q2
```

```{r}
contribution_data_all_states %>%
  group_by(Politician) %>%
  summarize(count_polititians = n()) %>%
  arrange(desc(count_polititians))
```

```{r}
#| label: years_greg_abbot

contribution_data_all_states %>%
  filter(Politician == "Greg Abbott") %>% 
  mutate(year = year(ymd(Date))) %>%
  select(year) %>%
  distinct(year)
```