---
title: "The Modern State of Pride and Anti-LGBTQ+ in the US"
subtitle: "STA/ISS 313 - Project 1"
author: "Co-Medians - Ellen Zhang, Amber Potter, Helen Chen, Morgan Bernstein"
format: html
editor: visual
---

## Abstract

(1 paragraph): Project abstract.

We are all advocates for LGBTQ+ rights, and we are very intrigued by the prevalence of pride supporters who also donate to anti-LGBTQ+ campaigns. While there has been growing support for the LGBTQ+ community in the US and increased discussion on human rights, diversity, and inclusion, there is still much more progress needed in this area. While many companies have publicly opposed anti-LGBTQ+ bills and have pro-LGBTQ campaigns, some of these companies simultaneously fund lawmakers who back anti-LGBTQ+ bills (The 19th). In particular, from 2021-2022, the newspaper "Popular Information" found that 25 companies who were publicly in support of the pride movement have also donated \$13M to politicians who back anti-LGBTQ+ bills. We believe this is dishonest behavior that deserves attention so more people are aware of these hypocritical practices. Further, politicians are undoubtedly vital to policy reform to support the LGBTQ+ community, however, there has been a significant amount of funding going towards politicians who are anti-LGBTQ+. If we want positive and real change to uplift the LGBTQ+ community, we need greater governmental support for pride and less for anti-LGBTQ+. In summary, for this project, we wanted to delve into two key players in the pride and anti-LGBTQ+ movements: companies and politicians, and understand how both stakeholder groups are associated with anti-LGBTQ+ campaigns that undermine progress to support pride in the US.

## Introduction

(1-2 paragraphs): Brief introduction to the dataset. You may repeat some of the information about the dataset provided in the introduction to the dataset on the TidyTuesday repository, paraphrasing on your own terms. Imagine that your project is a standalone document and the grader has no prior knowledge of the dataset.

```{r load-packages-for-intro, echo=FALSE, message=FALSE}
library(tidyverse)

# read data
static_list <- readRDS("data/static_list.RDS")

contribution_data_all_states <- readRDS("data/contribution_data_all_states.RDS")

corp_by_politician <-
  readr::read_csv(
    "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-06-07/corp_by_politician.csv"
  )
```

The datasets we are using are from [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-06-07) and come from [Data For Progress](https://www.dataforprogress.org/accountable-allie), which is a think tank and polling firm that provides data for progressive campaigns, causes, and candidates. Their work has been used by US representatives, senators, and local leaders, and it has been cited by the White House. The datasets in this TidyTuesday are looking specifically at businesses that contribute to the campaigns of politicians who have anti-LGBTQ beliefs, and it includes seven datasets: `static_list`, `contribution_data_all_states`, `pride_aggregates`, `fortune_aggregates`, `pride_sponsors`, `donors`, and `corp_by_politician`. We found overlapping data in `static_list`, `fortune_aggregates`, and `pride_aggregates`. `Static_list` captured the most complete information from `fortune_aggregates` and `pride_aggregates` combined, so we decided to use `static_list`. Our decision to use `contribution_data_all_states` was made for similar reasons. In many ways it is a combination of the others in terms of information and variables, but we appreciated that it included that information by individual contribution rather than by company or politician totals. Finally, we also decided to use the `corp_by_politician` dataset to have information on the offices and titles each politician holds.

The first dataset we are using is `static_list`, containing information on `r nrow(static_list)` companies (each row corresponds to a company) and whether they donated to pride, if they have an HRC business pledge, the dollar amount contributed across states (USD), the number of politicians contributed to, and the number of states where they made contributions. It has `r ncol(static_list)` columns/variables. A few of the variables we are using from this dataset for our analysis include `Company`, `Pride?`, `HRC Business Pledge`, and `Amount Contributed Across States`. The second dataset is `contribution_data_all_states` with `r nrow(contribution_data_all_states)` observations with each row denoting a single donation/contribution along with information on the company it was made by, whether or not they were a pride and sponsor match, if they were a pride event sponsor, HRC business pledge, donor name, politician donated to, state donated to, amount of donation (USD), date, citation, donor type, comments, and an archive. It has a total of `r ncol(contribution_data_all_states)` columns/variables. Some of the variables we are using to answer our research questions from this dataset are `Politician`, `Date`, and `Amount`. The last dataset is `corp_by_poltician`, which has information on `r nrow(corp_by_politician)` politicians and a total of `r ncol(corp_by_politician)` variables, including the politician's name, state, total amount of contributions received from business, and title. For this dataset, we are particularly interested in looking at politician `Title`.

## Question 1: Title that relates to the question you're answering

### Introduction

(1-2 paragraphs): Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you're interested in this question.

### Approach

(1-2 paragraphs): Describe what types of plots you are going to make to address your question. For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed): In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides. You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so. All plots must be made with **ggplot2**. Do not use base R or lattice plotting functions.

#### Plot 1:

```{r load-packages, echo=FALSE}
library(tidyverse)
library(dplyr)
library(tidytuesdayR)
library(lubridate)
library(readr)
library(ggpattern)
library(scales)

# read data
static_list <- readRDS("data/static_list.RDS")
contribution_data_all_states <- readRDS("data/contribution_data_all_states.RDS")
```

```{r data-processing-plot-1, echo=FALSE}
#| label: data_processing_plot_1

static_list <- static_list %>%
  mutate(
    amount_contributed = `Amount Contributed Across States`,
    hrc_pledge = `HRC Business Pledge`,
    pride_or_not = `Pride?`
  )

new_static_list <- static_list %>%
  filter(!is.na(Company), !is.na(hrc_pledge)) %>%
  select(Company, amount_contributed, hrc_pledge, pride_or_not) %>%
  group_by(Company) %>%
  summarise(
    total_amount = sum(amount_contributed),
    hrc_pledge, pride_or_not
  ) %>%
  arrange(desc(total_amount)) %>%
  slice_head(n = 20)
```

```{r question1-plot-1, echo=FALSE}
#| label: question 1 plot 1

new_static_list %>%
  mutate(
    hrc_pledge = if_else(hrc_pledge == FALSE, "No", "Yes"),
    pride_or_not = if_else(pride_or_not == FALSE, "No", "Yes")
  ) %>%
  ggplot(aes(x = Company, y = total_amount, fill = pride_or_not, pattern = hrc_pledge)) +
  geom_col_pattern() +
  scale_fill_manual(values = c("#AA4A44", "lightgrey")) +
  scale_pattern_manual(values = c("none", "stripe")) +
  scale_y_continuous(label = scales::label_dollar()) +
  coord_flip() +
  labs(
    title = "USD Amounts of top 20 companies in contributions \nto anti-LGBTQ+ campaigns",
    x = "Company Name",
    y = "Amount contributed to anti-LGBTQ+ campaigns (in USD)",
    fill = "Does the company \ndonate to Pride?",
    pattern = "Does the company \nhave an HRC pledge?"
  ) +
  theme_minimal() +
  theme(
    legend.position = c(.85, .32),
    legend.title = element_text(size = 8),
    legend.background = element_rect(fill = "white", color = "darkgrey")
  )
```

#### Plot 2:

```{r question1-plot-2, echo=FALSE, warning=FALSE}
#| label: question 1 plot 2
#| warning: false

static_list %>%
  mutate(
    category = case_when(
      hrc_pledge == TRUE & pride_or_not == TRUE ~ "HRC Pledge, donates to Pride",
      hrc_pledge == FALSE & pride_or_not == TRUE ~ "No HRC pledge, donates to Pride",
      TRUE ~ "Neither"
    ),
    category = fct_relevel(category, levels = c(
      "HRC Pledge, donates to Pride",
      "No HRC pledge, donates to Pride", "Neither"
    ))
  ) %>%
  ggplot(aes(x = amount_contributed)) +
  geom_density(aes(fill = category, color = category), show.legend = FALSE) +
  scale_x_continuous(label = scales::label_dollar()) +
  scale_y_continuous(labels = label_number()) + # idk if scientific notation is better
  scale_fill_manual(values = c("#AA4A44", "#BF8F8C", "lightgrey")) +
  scale_color_manual(values = c("#AA4A44", "#BF8F8C", "lightgrey")) + # can take this out
  facet_wrap(~category, ncol = 1) +
  coord_cartesian(xlim = c(0, 750000)) +
  labs(
    title = "Density of company contributions (USD) to anti-LGBTQ+ campaigns",
    subtitle = "Faceted by company characteristics", # CHANGE TITLE
    x = "Amount contributed (USD)",
    y = "Density"
  ) +
  theme_minimal()
```

### Discussion

(1-3 paragraphs): In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots. Speculate about why the data looks the way it does.

## Question 2: Title that relates to the question you're answering

### Introduction

(1-2 paragraphs): Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you're interested in this question.

Our second question is: how have donations to politicians who support anti-LGBTQ+ campaigns changed over time? We are very curious about how the total amount of money contributed to anti-LGBTQ+ politicians has changed over the years, especially given the growing discussions around human rights in the past few years. However, we also know that outspoken homophobia and transphobia seemed to grow during and following Trump's presidency.

In particular, Greg Abbott, the Governor of Texas, has been receiving a lot of public attention recently as an outspoken Republican. On Feb. 13, 2023, an article by the Human Rights Campaign denounced Abbott who said that he'll aim to ban transgender college athletes from playing any sports that are consistent with their gender identity. According to Time, in 2022, he also had an Anti-Trans Directive which negatively affected youth from the LGBTQ+ community; he said that the provision of gender-affirming care for children was a form a child abuse, and families were being investigated for this.

From our EDA, we also realized that Abbott received the highest number of individual donations overall, despite him only being included over the years 2019-2022. As an extension of our main question, we are interested in looking into donations that were made specifically to Greg Abbott and seeing how his amounts compare to overall trends across all politicians. Overall, we would hypothesize that the trend of anti-LGBTQ+ political campaign donations would decrease over time (or we would hope they do), however, due to the large media presence of Greg Abbott and his strong rhetoric on this topic, we believe he might prove a contrast to the overall trend.

### Approach

(1-2 paragraphs): Describe what types of plots you are going to make to address your question. For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

We were interested in how the total USD amounts contributed to politicians changed over time. To do this, we created a stacked bar plot with year on the x axis and total amount (USD) on the y axis. The height of each bar indicated that year's total money received by anti-LGBTQ+ politicians across the states included. Diving further, we wanted to also look at Greg Abbott and how he compares to the overall amount. For the stacked bar plot, we colored by Greg Abbott vs. all the other politicians except Abbott. Since Abbott only had data for the years 2019-2022, we focused on those four years in our plot.

We understand that a challenge with stacked bar plots is that it is hard to compare them when their heights are different. So, we added annotations for the different proportions (in %) to make the stacked bar plots easier to interpret and to make it more clear how the contributions to Abbott compared to those of the other politicians. We found the stacked bar plot to be the most appropriate for this question because it shows how the total donation amounts were divided between politicians who were and were not Greg Abbott. It also clearly visualizes the relationship of each group of politician on the overall donation amount. Overall, the plot we created was helpful in comparing the total donation amounts for politicians who were and were not Greg Abbott simultaneously and understand the different proportions in donations over the years.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed): In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides. You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so. All plots must be made with **ggplot2**. Do not use base R or lattice plotting functions.

#### Plot 4:

```{r load-merge-and-process-data, echo=FALSE}
corp_by_politician <-
  readr::read_csv(
    "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-06-07/corp_by_politician.csv"
  )
static_list <- readRDS("data/static_list.RDS")
contribution_data_all_states <-
  readRDS("data/contribution_data_all_states.RDS")

question2_data <-
  inner_join(contribution_data_all_states, corp_by_politician, by = "Politician")
# inner join to capture the politicians that appear in both datasets

question2_plot2_data <- question2_data %>%
  filter(!is.na(Date)) %>%
  mutate(year = year(ymd(Date))) %>%
  mutate(is_greg = ifelse(Politician == "Greg Abbott", "Greg Abbott", "Not Greg Abbott")) %>%
  filter(year == 2019 | year == 2020 | year == 2021 | year == 2022)

# question2_data %>% filter(Politician == "Greg Abbott") %>% select(year) %>% unique()
# Greg only had data in years 2019-2022 so we consider only those years
```

```{r plot-4-more-data-processing-for-plot, echo=FALSE}
greg_not_greg <- question2_plot2_data %>%
  group_by(is_greg, year) %>%
  summarise(
    amounts = sum(Amount),
    .groups = "drop"
  )

not_greg <- greg_not_greg %>%
  filter(is_greg == "Not Greg Abbott")

props_greg_not_greg <- greg_not_greg %>%
  group_by(year) %>%
  summarise(
    prop = round(100 * (amounts / sum(amounts)), 0),
    .groups = "drop"
  )

total_amount_per_year <- greg_not_greg %>%
  group_by(year) %>%
  summarise(
    amount_sum = sum(amounts),
    .groups = "drop"
  )

annotation_data_2019 <- tribble(
  ~amount_min,
  ~amount_max,
  ~percentage,
  ~color,
  0,
  not_greg$amounts[1],
  paste0(props_greg_not_greg$prop[2], "%"),
  "black",
  not_greg$amounts[1],
  total_amount_per_year$amount_sum[1],
  paste0(props_greg_not_greg$prop[1], "%"),
  "white"
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2020 <- tribble(
  ~amount_min,
  ~amount_max,
  ~percentage,
  ~color,
  0,
  not_greg$amounts[2],
  paste0(props_greg_not_greg$prop[4], "%"),
  "black",
  not_greg$amounts[2],
  total_amount_per_year$amount_sum[2],
  paste0(props_greg_not_greg$prop[3], "%"),
  "white"
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2021 <- tribble(
  ~amount_min,
  ~amount_max,
  ~percentage,
  ~color,
  0,
  not_greg$amounts[3],
  paste0(props_greg_not_greg$prop[6], "%"),
  "black",
  not_greg$amounts[3],
  total_amount_per_year$amount_sum[3],
  paste0(props_greg_not_greg$prop[5], "%"),
  "white"
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2022 <- tribble(
  ~amount_min,
  ~amount_max,
  ~percentage,
  ~color,
  0,
  not_greg$amounts[4],
  paste0(props_greg_not_greg$prop[8], "%"),
  "black",
  not_greg$amounts[4],
  total_amount_per_year$amount_sum[4],
  paste0(props_greg_not_greg$prop[7], "%"),
  "white"
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)
```

```{r plot-4, echo=FALSE}
ggplot(greg_not_greg) +
  geom_bar(aes(x = year, y = amounts, fill = is_greg),
    position = "stack",
    stat = "identity"
  ) +
  geom_text(
    data = annotation_data_2019,
    aes(
      x = 2019,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2020,
    aes(
      x = 2020,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2021,
    aes(
      x = 2021,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2022,
    aes(
      x = 2022,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  scale_color_identity() +
  labs(
    x = "Year", y = "Amount Donated for anti-LGBTQ+ Campaigns
       (in USD)",
    fill = "Politician",
    title = "Distribution of donation amounts for anti-LGBTQ+ politicians",
    subtitle = "In the years 2019-2022"
  ) +
  scale_fill_manual(values = c("#AA4A44", "lightgrey")) +
  scale_y_continuous(
    limits = c(0, 11000000),
    labels = scales::label_dollar(),
    breaks = c(0, 2000000, 4000000, 6000000, 8000000, 10000000)
  ) +
  theme_minimal() +
  theme(
    legend.position = c(0.14, 0.85),
    legend.background = element_blank(),
    legend.box.background = element_rect(color = "darkgrey", size = 0.5)
  )
```

### Discussion

(1-3 paragraphs): In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots. Speculate about why the data looks the way it does.

From our fourth plot, we can see that in the years 2019-2021, there has been an increase in the amounts donated to politicians who are anti-LGBTQ+, with a stark increase in the year 2021. Upon further research, we found that 2021 was the worst year for anti-LGBTQ+ legislation since 2015 (HRC). In 2021, there were 17 anti-LGBTQ bills that were enacted into law, and this was more than the number of anti-LGBTQ+ laws in the years 2018-2020 combined. Some of these include bills that ban transgender girls from being involved in sports to the erasure of LGBTQ+ individuals from curriculum in schools to felonies for giving transgender young people healthcare in life-saving situations to licenses to discriminate against those in the LGBTQ+ community. In fact, there were 250+ bills that were introduced in legislatures that year. Fortunately, the total amounts decreased significantly in the year 2022, and we hope this downward trend will continue. Additionally, we can see that the proportion of total donations going to Greg Abbott increased in the years 2019-2021, with him having over 60% of total donations in 2021; not only have the proportions increased during those years, but the dollar amounts he received also increased. This suggests that there was, unfortunately, a lot of support for Abbott, as he has always been very vocal about being anti-LGBTQ+ and is a prominent figure in the US, being the Governor of Texas, a traditionally very red state. Fortunately, he received a smaller proportion of the total donations in 2022 and a smaller dollar amount of donations. This could be associated with the fact that he got the worst press out of all governors in the US regarding LGBTQ+ issues (Houston Chronicle), and this could've led less people to want to donate to his controversial campaigns.

A limitation of our plot is that it only includes information on six states: Florida, Texas, Idaho, Alabama, Arizona, and Tennessee, and there are many more states that are anti-LGBTQ+ that we weren't able to capture due to limited data; for instance, according to NBC News, 23 states don't protect citizens from discrimination on the basis of sexual orientation. The six states that were in this dataset were some of the most prominent and alarming anti-LGBTQ+ supporters, however, our research would've been stronger had there been information on more states. For instance, if the dataset included blue states, we could've compared the difference (if any) in support for anti-LGBTQ between Republican and Democratic politicians.


#### Plot 3:

```{r}
question2_data <-
  inner_join(contribution_data_all_states, corp_by_politician, by = "Politician")

question2_plot3_data <- question2_data %>%
  filter(!is.na(Date)) %>%
  mutate(year = year(ymd(Date))) %>%
  #mutate(is_greg = ifelse(Politician == "Greg Abbott", "Greg Abbott", "Not Greg Abbott")) %>%
  group_by(Politician, year, Title) %>%
  summarize(avg_amount = mean(Amount, na.rm = TRUE))

plot3_data_no_abbott <- question2_plot3_data %>%
  filter(Politician != "Greg Abbott") %>%
  mutate(position_cat = case_when(Title == "Representative" ~ "Representative",
                                  Title == "Senator" ~ "Senator",
                                  Title == "Governor" ~ "Governor",
                                  TRUE ~ "Other")) %>%
  group_by(position_cat, year) %>%
  summarize(avg_avg_amount = mean(avg_amount, na.rm = TRUE))
  

plot3_data_abbott <- question2_plot3_data %>%
  filter(Politician == "Greg Abbott")

unique(plot3_data_no_abbott$position_cat)


ggplot() +
  geom_line(data = plot3_data_no_abbott, mapping = aes(x = year, y = avg_avg_amount, color = position_cat)) +
  geom_line(plot3_data_abbott, mapping = aes(x = year, y = avg_amount))
```


## Sources:

-   https://19thnews.org/2022/06/companies-anti-lgbtq-bills-lawmakers/

-   https://www.hrc.org/press-releases/human-rights-campaign-gov-abbott-continuing-his-crusade-to-harm-transgender-and-non-binary-texans

-   https://time.com/6150964/greg-abbott-trans-kids-child-abuse/

-   https://www.hrc.org/press-releases/2021-officially-becomes-worst-year-in-recent-history-for-lgbtq-state-legislative-attacks-as-unprecedented-number-of-states-enact-record-shattering-number-of-anti-lgbtq-measures-into-law

-   https://www.houstonchronicle.com/politics/texas/article/gov-greg-gbbott-gets-worst-press-on-lgbtq-issues-17184157.php

-   https://www.nbcnews.com/nbc-out/out-politics-and-policy/less-10-2022s-anti-lgbtq-state-bills-became-law-report-finds-rcna67619

-   https://meyerweb.com/eric/tools/color-blend/#D3D3D3:AA4A44:1:hex
