---
title: "Project title"
subtitle: "STA/ISS 313 - Project 1"
author: "Team Name"
format: html
editor: visual
---

## Abstract

(1 paragraph): Project abstract.

------------------------------------------------------------------------

## Introduction

(1-2 paragraphs): Brief introduction to the dataset. You may repeat some of the information about the dataset provided in the introduction to the dataset on the TidyTuesday repository, paraphrasing on your own terms. Imagine that your project is a standalone document and the grader has no prior knowledge of the dataset.

## Question 1: Title that relates to the question you're answering

### Introduction

(1-2 paragraphs): Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you're interested in this question.

### Approach

(1-2 paragraphs): Describe what types of plots you are going to make to address your question. For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed): In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides. You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so. All plots must be made with **ggplot2**. Do not use base R or lattice plotting functions.

```{r load-packages}
library(tidyverse)
library(lubridate)
```

```{r load-data}
corp_by_politician <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-06-07/corp_by_politician.csv')
static_list <-readRDS("data/static_list.RDS")
contribution_data_all_states <-readRDS("data/contribution_data_all_states.RDS")

question2_data <- inner_join(contribution_data_all_states, corp_by_politician, by = "Politician")

question2_data <- question2_data %>%
  filter(!is.na(Date)) %>%
  mutate(year = year(ymd(Date))) %>%
  mutate(is_greg = ifelse(Politician == "Greg Abbott", "Greg Abbott", "Not Greg")) %>%
  filter(year == 2019 | year == 2020 | year == 2021 | year == 2022) 

# question2_data %>% filter(Politician == "Greg Abbott") %>% select(year) %>% unique()
# Greg only had data in years 2019-2022 so we consider only those years
```

```{r plot-4-data-processing}
greg_not_greg <- question2_data %>%
  group_by(is_greg, year) %>%
  summarise(amounts = sum(Amount),
            .groups = 'drop')

not_greg <- greg_not_greg %>%
  filter(is_greg == "Not Greg")

props_greg_not_greg <- greg_not_greg %>%
  group_by(year) %>%
  summarise(prop = round(100 * (amounts / sum(amounts)), 0),
            .groups = 'drop')

total_amount_per_year <- greg_not_greg %>%
  group_by(year) %>%
  summarise(amount_sum = sum(amounts), 
            .groups = 'drop')

annotation_data_2019 <- tribble(
  ~ amount_min,
  ~ amount_max,
  ~ percentage,
  ~ color, 
  0,
  not_greg$amounts[1],
  paste0(props_greg_not_greg$prop[props_greg_not_greg$year == 2019][2], "%"),
  'black',
  not_greg$amounts[1],
  total_amount_per_year$amount_sum[1],
  paste0(props_greg_not_greg$prop[props_greg_not_greg$year == 2019][1], "%"),
  'white'
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2020 <- tribble(
  ~ amount_min,
  ~ amount_max,
  ~ percentage,
  ~ color,
  0,
  not_greg$amounts[2],
  paste0(props_greg_not_greg$prop[props_greg_not_greg$year == 2020][2], "%"),
  'black',
  not_greg$amounts[2],
  total_amount_per_year$amount_sum[2],
  paste0(props_greg_not_greg$prop[props_greg_not_greg$year == 2020][1], "%"),
  'white'
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2021 <- tribble(
  ~ amount_min,
  ~ amount_max,
  ~ percentage,
  ~ color, 
  0,
  not_greg$amounts[3],
  paste0(props_greg_not_greg$prop[props_greg_not_greg$year == 2021][2], "%"),  
  'black',
  not_greg$amounts[3],
  total_amount_per_year$amount_sum[3],
  paste0(props_greg_not_greg$prop[props_greg_not_greg$year == 2021][1], "%"),
  'white'
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2022 <- tribble(
  ~ amount_min,
  ~ amount_max,
  ~ percentage,
  ~ color,
  0,
  not_greg$amounts[4],
  paste0(props_greg_not_greg$prop[props_greg_not_greg$year == 2022][2], "%"),
  'black',
  not_greg$amounts[4],
  total_amount_per_year$amount_sum[4],
  paste0(props_greg_not_greg$prop[props_greg_not_greg$year == 2022][1], "%"),
  'white'
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)
```

```{r plot-4}
ggplot(greg_not_greg) +
  geom_bar(aes(x = year, y = amounts, fill = is_greg),
           position = 'stack',
           stat = 'identity') +
  geom_text(
    data = annotation_data_2019,
    aes(
      x = 2019,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2020,
    aes(
      x = 2020,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2021,
    aes(
      x = 2021,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2022,
    aes(
      x = 2022,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  scale_color_identity() +
  labs(x = "Year", y = "Amount Donated for anti-LGBTQ+ Campaigns 
       (in USD)",
       fill = "Politician",
       title = "Distribution of donation amounts for anti-LGBTQ+ campaigns to politicians",
       subtitle = "In the years 2019-2022") +
  scale_fill_manual(values = c("#AA4A44", "lightgrey")) +
  scale_y_continuous(
    limits = c(0, 11000000),
    labels = scales::label_dollar(),
    breaks = c(0, 2000000, 4000000, 6000000, 8000000, 10000000)
  ) +
  theme_minimal() +
   theme(
    legend.position = c(0.14, 0.85),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "darkgrey", size = 0.5))
```

### Discussion

(1-3 paragraphs): In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots. Speculate about why the data looks the way it does.

## Question 2: Title that relates to the question you're answering

### Introduction

(1-2 paragraphs): Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you're interested in this question.

Our second question is: how have donations to politicians who support anti-LGBTQ+ campaigns changed over time? We are very curious about how the total amount of money contributed to anti-LGBTQ+ politicians has changed over the years, especially given the growing discussions around human rights in the past few years. However, we also know that outspoken homophobia and transphobia seemed to grow during and following Trump's presidency.

In particular, Greg Abbott, the Governor of Texas, has been receiving a lot of public attention recently as an outspoken Republican. On Feb 13, 2023, an article by the Human Rights Campaign noted that Abbott said he'll aim to ban transgender college athletes from playing any sports that are consistent with their gender identity (https://www.hrc.org/press-releases/human-rights-campaign-gov-abbott-continuing-his-crusade-to-harm-transgender-and-non-binary-texans). According to Time, in 2022, he also had an Anti-Trans Directive which negatively affected youth from the LGBTQ+ community (https://time.com/6150964/greg-abbott-trans-kids-child-abuse/). From our EDA, we realized that Abbott received the highest number of individual donations overall, despite him only being included over the years 2019-2022. As an extension of our main question, we are also interested in looking into donations that were made specifically to Greg Abbott and seeing how his amounts compare to overall trends across all politicians. Overall, we would hypothesize that the trend of anti-LGBTQ+ political campaign donations would decrease over time (or one would hope they do), however, due to the large media presence of Greg Abbott and his strong rhetoric involving this topic, we believe he might prove a contrast to the overall trend.

### Approach

(1-2 paragraphs): Describe what types of plots you are going to make to address your question. For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed): In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides. You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so. All plots must be made with **ggplot2**. Do not use base R or lattice plotting functions.

### Discussion

(1-3 paragraphs): In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots. Speculate about why the data looks the way it does.
