---
title: "The Hypocrisy of US Corporate Pride and Anti-LGBTQ+ Donations"
format: revealjs
editor: visual
---

```{r load-packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(tidytuesdayR)
library(lubridate)
library(readr)
library(ggpattern)
library(scales)

# read data
static_list <- readRDS("data/static_list.RDS")

contribution_data_all_states <- readRDS("data/contribution_data_all_states.RDS")

corp_by_politician <-
  readr::read_csv(
    "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-06-07/corp_by_politician.csv"
  )
```

```{r bullets, echo=FALSE}
# Source for our bullet structure in this presentation: https://stackoverflow.com/questions/49084904/r-markdown-bullet-list-with-multiple-levels
```

## Data {.smaller}

-   Datasets are from TidyTuesday and originally drawn from Data for Progress

-   Data looks specifically at businesses that contribute to anti-LGBTQ political campaigns

-   We use 3 out of the 7 total datasets:

    -   Static_list - most complete combined info from other datasets

    -   Contribution_data_all_states - info by individual contribution

    -   corp_by_politician - info on politician offices

## Question 1: What is the relationship between a company's status as a pride sponsor, whether they have an HRC pledge or not, and their contributions to anti-LGBTQ+ political campaigns?

::: aside
\[\*\] The Human Rights Campaign (HRC) business pledge is a commitment for companies to show their clear opposition to harmful legislation aimed at restricting the access of LGBTQ+ identifying people in society and the workplace, and 296 companies have already signed the pledge
:::

## Q1 Plot 1

```{r data-processing-plot-1, echo=FALSE}
#| label: data_processing_plot_1

static_list <- static_list %>%
  mutate(
    amount_contributed = `Amount Contributed Across States`,
    hrc_pledge = `HRC Business Pledge`,
    pride_or_not = `Pride?`
  )

new_static_list <- static_list %>%
  filter(!is.na(Company), !is.na(hrc_pledge)) %>%
  select(Company, amount_contributed, hrc_pledge, pride_or_not) %>%
  group_by(Company) %>%
  summarise(
    total_amount = sum(amount_contributed),
    hrc_pledge, pride_or_not
  ) %>%
  arrange(desc(total_amount)) %>%
  slice_head(n = 20)
```

```{r question1-plot-1, echo=FALSE}
#| label: question 1 plot 1

new_static_list %>%
  mutate(
    hrc_pledge = if_else(hrc_pledge == FALSE, "No", "Yes"),
    pride_or_not = if_else(pride_or_not == FALSE, "No", "Yes")
  ) %>%
  ggplot(aes(x = reorder(Company,total_amount), y = total_amount, fill = pride_or_not, pattern = hrc_pledge)) +
  geom_col_pattern(pattern_fill = "darkgray", pattern_color = "darkgray", aes(color = pride_or_not), width = .7) +
  scale_color_manual(values = c("#AA4A44", "lightgrey"), guide = "none") +
  scale_fill_manual(values = c("#AA4A44", "lightgrey")) +
  scale_pattern_manual(values = c("none", "stripe"), guide = guide_legend(override.aes = list(fill = "white", color = "darkgray"))) +
  scale_y_continuous(label = scales::label_dollar()) +
  coord_flip() +
  labs(
    title = "USD Amounts of top 20 companies in contributions \nto anti-LGBTQ+ campaigns",
    x = "Company",
    y = "Amount contributed to anti-LGBTQ+ campaigns (in USD)",
    fill = "Does the company \ndonate to Pride?",
    pattern = "Does the company \nhave an HRC pledge?"
  ) +
  theme_minimal() +
  theme(
    legend.position = c(.85, .32),
    legend.title = element_text(size = 8),
    legend.background = element_rect(fill = "white", color = "darkgrey")
  )
```

## Q1 Plot 1 Approach

-   **Goal:** Investigate the relationship between a company's contributions to anti-LGBTQ+ campaigns, their position as a Pride sponsor, and implementation of a HRC pledge

-   **Methodology:** Bar chart with amount contributed on the x-axis and company name on the y-axis

    -   Colored by whether or not the company is a Pride sponsor
    -   Striped bars indicate that the company has enacted an HRC pledge

## Q1 Plot 1 Discussion

-   Most of the top 20 companies have total contributions under \$100,000, with the exception of large outliers such as Toyota and AT&T
-   12 out of the 20 top companies are not Pride sponsors, however the top two companies in USD contributions, Toyota and AT&T, both contribute to Pride
-   The majority of companies in the top 20 also have not implemented an HRC pledge

## Q1 Plot 1 Limitations & Future Study {.smaller}

-   Dataset only includes information on contributions to politicians from six states: Florida, Texas, Idaho, Alabama, Arizona, and Tennessee

-   Our investigation could be furthered by looking into the magnitude of donations to Pride vs to anti-LGBTQ+ campaigns to see if these companies are contributing equally to both or not

-   It would be interesting to look at company efforts to promote exclusivity and equality in the workplace other than an HRC pledge, like on their company website

## Q1 Plot 2

```{r question1-plot-2-data, echo=FALSE}
company_categ_counts <- static_list %>%
  mutate(
    category = case_when(
      hrc_pledge == TRUE &
        pride_or_not == TRUE ~ "HRC Pledge, donates to Pride",
      hrc_pledge == FALSE &
        pride_or_not == TRUE ~ "No HRC pledge, donates to Pride",
      hrc_pledge == TRUE &
        pride_or_not == FALSE ~ "HRC Pledge, no donation to Pride",
      TRUE ~ "Neither"
    ),
    category = fct_relevel(factor(
      category,
      levels = c(
        "HRC Pledge, donates to Pride",
        "No HRC pledge, donates to Pride",
        "HRC Pledge, no donation to Pride",
        "Neither"
      )
    ))
  ) %>%
  count(category)
```

```{r question1-plot-2, echo=FALSE, warning=FALSE}
#| label: question 1 plot 2
#| warning: false

static_list %>%
  mutate(
    category = case_when(
      hrc_pledge == TRUE &
        pride_or_not == TRUE ~ "HRC Pledge, donates to Pride",
      hrc_pledge == FALSE &
        pride_or_not == TRUE ~ "No HRC pledge, donates to Pride",
      hrc_pledge == TRUE &
        pride_or_not == FALSE ~ "HRC Pledge, no donation to Pride",
      TRUE ~ "Neither"
    ),
    category = fct_relevel(factor(
      category,
      levels = c(
        "HRC Pledge, donates to Pride",
        "No HRC pledge, donates to Pride",
        "HRC Pledge, no donation to Pride",
        "Neither"
      )
    ))
  ) %>%
  ggplot(aes(x = amount_contributed)) +
  geom_density(aes(fill = category, color = category), show.legend = FALSE) +
  scale_x_continuous(label = scales::label_dollar()) +
  scale_y_continuous(labels = label_number()) + # idk if scientific notation is better
  scale_fill_manual(values = c("#AA4A44", "#BF8F8C", "grey", "#899499")) +
  scale_color_manual(values = c("#AA4A44", "#BF8F8C", "grey", "#899499")) + # can take this out
  facet_wrap(~ category, ncol = 1) +
  coord_cartesian(xlim = c(0, 700000)) +
  labs(
    title = "Density of company contributions to anti-LGBTQ+ campaigns",
    subtitle = "Faceted by company characteristics",
    # CHANGE TITLE
    x = "Amount contributed (in USD)",
    y = "Density"
  ) +
  theme_minimal() +
  geom_text(data = company_categ_counts,
            aes(
              x = 680000,
              y = 0.000004,
              label = paste0("n = ", company_categ_counts$n),
            ),
            size = 2.6)
```

## Q1 Plot 2 Approach {.smaller}

-   Goal: Investigate the relationship between a company's contributions to anti-LGBTQ+ campaigns, position as a Pride sponsor, and existence of an HRC pledge.

-   Methodology: Density plot showing the density of the number of contributions

    -   Colored & faceted by category of company: 1) HRC pledge and donates to Pride, 2) HRC pledge and doesn't donate to Pride, 3) No HRC pledge and donates to Pride, 4) Neither HRC pledge nor donation to Pride
    -   Annotations denoting the number of companies that fall into each category

## Q1 Plot 2 Discussion

## Q1 Plot 2 Limitations & Future Study

-   Numbers of companies were different across each group therefore drawing meaningful conclusions is more difficult

-   In addition to this the range of sample sizes is quite large, with many companies falling into the 'neither' category, which is about 4 times as large as the second largest category

## Question 2: How have donations to politicians who support anti-LGBTQ+ campaigns changed over time?

## Q2 Plot 1

## Q2 Plot 1 Approach

-   Goal: Visualize the fluctuations in average contribution per politician over time, specifically Greg Abbott compared with other politicians

-   Methodology: Line plot with year on the x-axis and average contribution per politician per year on the y-axis

    -   Colored by politician title
    -   Dashed line to indicate Greg Abbott
    -   Annotations denoting which group each line represents

## Q2 Plot 1 Discussion

## Q2 Plot 1 Limitations & Future Study

-   Greg Abbott is only present from the years 2019-2022 which greatly limits our analysis

-   There was only one other politician with the title governor where we had information on contribution dates. It would be great to clearly visualize Greg Abbott's average contributions compared with others of his political influence

-   By displaying the average contribution amount per politician we lose specific information on the number of donations that were made as well as the magnitude of donations

## Q2 Plot 2

```{r load-merge-and-process-data-for-plot-2, echo=FALSE, warning=FALSE}
corp_by_politician <-
  readr::read_csv(
    "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-06-07/corp_by_politician.csv"
  )
static_list <- readRDS("data/static_list.RDS")
contribution_data_all_states <-
  readRDS("data/contribution_data_all_states.RDS")

question2_data <-
  inner_join(contribution_data_all_states, corp_by_politician, by = "Politician")
# inner join to capture the politicians that appear in both datasets

question2_plot2_data <- question2_data %>%
  filter(!is.na(Date)) %>%
  mutate(year = year(ymd(Date))) %>%
  mutate(is_greg = ifelse(Politician == "Greg Abbott", "Greg Abbott", "Not Greg Abbott")) %>%
  filter(year == 2019 | year == 2020 | year == 2021 | year == 2022)

# question2_data %>% filter(Politician == "Greg Abbott") %>% select(year) %>% unique()
# Greg only had data in years 2019-2022 so we consider only those years
```

```{r plot-4-more-data-processing-for-plot, echo=FALSE, warning=FALSE}
greg_not_greg <- question2_plot2_data %>%
  group_by(is_greg, year) %>%
  summarise(
    amounts = sum(Amount),
    .groups = "drop"
  )

greg <- greg_not_greg %>%
  filter(is_greg == "Greg Abbott")

props_greg_not_greg <- greg_not_greg %>%
  group_by(year) %>%
  summarise(
    prop = round(100 * (amounts / sum(amounts)), 0),
    .groups = "drop"
  )

total_amount_per_year <- greg_not_greg %>%
  group_by(year) %>%
  summarise(
    amount_sum = sum(amounts),
    .groups = "drop"
  )

annotation_data_2019 <- tribble(
  ~amount_min,
  ~amount_max,
  ~percentage,
  ~color,
  0,
  greg$amounts[1],
  paste0(props_greg_not_greg$prop[1], "%"),
  "white",
  greg$amounts[1],
  total_amount_per_year$amount_sum[1],
  paste0(props_greg_not_greg$prop[2], "%"),
  "black"
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2020 <- tribble(
  ~amount_min,
  ~amount_max,
  ~percentage,
  ~color,
  0,
  greg$amounts[2],
  paste0(props_greg_not_greg$prop[3], "%"),
  "white",
  greg$amounts[2],
  total_amount_per_year$amount_sum[2],
  paste0(props_greg_not_greg$prop[4], "%"),
  "black"
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2021 <- tribble(
  ~amount_min,
  ~amount_max,
  ~percentage,
  ~color,
  0,
  greg$amounts[3],
  paste0(props_greg_not_greg$prop[5], "%"),
  "white",
  greg$amounts[3],
  total_amount_per_year$amount_sum[3],
  paste0(props_greg_not_greg$prop[6], "%"),
  "black"
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)

annotation_data_2022 <- tribble(
  ~amount_min,
  ~amount_max,
  ~percentage,
  ~color,
  0,
  greg$amounts[4],
  paste0(props_greg_not_greg$prop[7], "%"),
  "white",
  greg$amounts[4],
  total_amount_per_year$amount_sum[4],
  paste0(props_greg_not_greg$prop[8], "%"),
  "black"
) %>%
  mutate(amount_mid = (amount_min + amount_max) / 2)
```

```{r plot-4, echo=FALSE, warning=FALSE}
ggplot(greg_not_greg) +
  geom_bar(aes(
    x = year,
    y = amounts,
    fill = fct_relevel(factor(
      is_greg, levels = c("Not Greg Abbott", "Greg Abbott")
    ))
  ),
  position = "stack",
  stat = "identity") +
  geom_text(
    data = annotation_data_2019,
    aes(
      x = 2019,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2020,
    aes(
      x = 2020,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2021,
    aes(
      x = 2021,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  geom_text(
    data = annotation_data_2022,
    aes(
      x = 2022,
      y = amount_mid,
      label = percentage,
      color = color
    ),
    size = 3
  ) +
  scale_color_identity() +
  labs(
    x = "Year", y = "Amount Donated for anti-LGBTQ+ Campaigns
       (in USD)",
    fill = "Politician",
    title = "Distribution of donation amounts for anti-LGBTQ+ politicians",
    subtitle = "In the years 2019-2022"
  ) +
  scale_fill_manual(values = c("lightgrey", "#AA4A44")) +
  scale_y_continuous(
    limits = c(0, 11000000),
    labels = scales::label_dollar(),
    breaks = c(0, 2000000, 4000000, 6000000, 8000000, 10000000)
  ) +
  theme_minimal() +
  theme(
    legend.position = c(0.12, 0.85),
    legend.background = element_blank(),
    legend.box.background = element_rect(color = "darkgrey", size = 0.5)
  )
```

## Q2 Plot 2 Approach

-   Goal: examine how the total amount contributed to anti-LGBTQ+ politicians changed over time

-   Methodology: stacked bar plot with year on the x axis and total amount of contributions on the y axis

    -   Colored by Greg Abbott vs. all the other politicians except Abbott
    -   Annotations for the different proportions (in %) for better interpret-ability

## Q2 Plot 2 Discussion

-   Increase in total amounts donated to anti-LGBTQ+ politicians in 2019-2021
    -   Stark increase in 2021; total amounts decreased significantly in 2022
-   The proportion (and dollar amount) of total donations going to Greg Abbott increased in 2019-2021
    -   Abbott had over 60% of total donations in 2021
-   In 2022, Abbott received a smaller proportion of total donations and a smaller dollar amount of donations

## Q2 Plot 2 Limitations & Future Study

-   Dataset only includes information on six states: Florida, Texas, Idaho, Alabama, Arizona, and Tennessee

-   Our research would've been stronger had there been information on more states so we could compare between Republican and Democratic politicians

    -   Would also be interesting to look into federal politicians

## Sources
