---
title: "Project 1: Co-Medians"
subtitle: "Project 1 Proposal"
authors: "Ellen Zhang, Amber Potter, Helen Chen, Morgan Bernstein"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
#| warning: false

library(tidyverse)
library(dplyr)
library(tidytuesdayR)
library(lubridate)
library(readr)
```

# Dataset

```{r}
#| label: load-data
#| message: false

# read data 
static_list <-readRDS("data/static_list.RDS")
contribution_data_all_states <-readRDS("data/contribution_data_all_states.RDS")
```

The datasets we are using are from TidyTuesday (https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-06-07) and come from the website Data For Progress. The datasets given were static_list, contribution_data_all_states, pride_aggregates, fortune_aggregates, pride_sponsors, donors, and corp_by_politician. After looking into the datasets further, we realized that there was overlap of information in static_list, fortune_aggregates, and pride_aggregates, where static_list captured the most complete information from fortune_aggregates and pride_aggregates combined, so we decided to use that dataset. We decided not to focus on specific pride events nor on politicians, so we didn't use the pride_sponsors nor corp_by_politician datasets. We also did not use information on specific donors, so that csv was disregarded as well. We did use the contribution_data_all_states dataset, as that provided date information (for use in time-series plots), contribution amount, and if the contribution was towards pride or anti-LGBTQ campaigns.

The first dataset we are using is static_list, containing information on `r nrow(static_list)` companies (each row corresponds to a company) and whether they donated to pride, HRC business pledge, dollar amount contributed across states, number of politicians contributed to, and number of states where they made contributions; there are `r ncol(static_list)` columns. A few of the variables we are using from this dataset for our analysis include `Pride?`, `HRC Business Pledge`, and `Amount Contributed Across States`. The second dataset is contribution_data_all_states with `r nrow(contribution_data_all_states)` observations with each row denoting a single donation, along with information on the company it was made by, whether or not they were a pride and sponsor match, if they were a pride event sponsor, HRC business pledge, donor name, politician donated to, state donated to, amount of donation, date, citation, donor type, comments, and an archive; there are a total of `r ncol(contribution_data_all_states)` columns. All contribution amounts are in USD. Some of the variables we are using to answer our research questions from this dataset are `Pride?`, `Date` (which we extracted the column `Year` from), and `Amount`.

# Reason for Choosing this Dataset

Overall, we thought that these datasets would be extremely interesting both to visualize and to learn about. We are all advocates for LGBTQ+ rights, and we were very intrigued by the prevalence of Fortune 500 pride sponsors who donate to anti-LGBTQ+ campaigns. We believe this is sneaky and dishonest behavior by corporations that deserves attention in order to help others become more aware of these hypocritical practices.

# Questions

## Question 1

What is the relationship between whether a company is a pride sponsor, has an hrc pledge, and the amount that they donate to anti-lgbt campaigns?

- plot 1: proportion of # of anti lgbt donations coming from hrc companies vs not
or proportion of anti lgbt donations coming from pride sponsors

- plot 2: comparing individual companies total donation amounts to anti lgbt, coloring by hrc

The first question we want to answer using this dataset is: which companies tend to donate the most money in contributions to anti-LGBTQ+ campaigns, which of these companies have a HRC pledge in place, and which of these companies have donated to pride campaigns? The Human Rights Campaign (HRC) business pledge is a commitment for companies to show their clear opposition to harmful legislation aimed at restricting the access of LGBTQ people in society and the workplace.

Source: https://www.hrc.org/resources/business-statement-on-anti-lgbtq-state-legislation

In our data processing, we renamed some variables (Amount Contributed Across States, HRC Business Pledge, and Pride?) for better readability.

```{r question-1-data-processing}

static_list <- static_list %>%
  rename(amount_contributed = `Amount Contributed Across States`)

static_list <- static_list %>%
  rename(hrc_pledge = `HRC Business Pledge`)

static_list <- static_list %>%
  rename(pride_or_not = `Pride?`)

new_static_list <- static_list %>%  
  filter(!is.na(Company), !is.na(hrc_pledge)) %>%
  select(Company, amount_contributed, hrc_pledge, pride_or_not) %>%
  group_by(Company) %>%
  summarise(total_amount = sum(amount_contributed),
            hrc_pledge, pride_or_not) %>%
  arrange(desc(total_amount)) %>%
  slice_head(n = 10)
  
new_static_list %>%
  filter(hrc_pledge == TRUE, pride_or_not == FALSE)
#idea for plot: x axis is contribution amount, y is the company, color whether or not have donated to pride, made stripes or another pattern if there's an HRC business pledge
```

## Question 2

Our second question involves the dollar amount of contributions made to anti-LGBTQ+ campaigns vs the dollar amount of contributions to pride campaigns by the hypocritical companies who contribute to both anti-LGBTQ+ and pro-pride campaigns. Specifically, what is the magnitude of difference in contributions to pro- and anti-LGBTQ+ campaigns made by these companies over the years 2013 to 2022, and are there any associations we can see between the two?

For our data processing, we created a year variable by extracting the year from the Date variable. We also filtered by companies that weren't NA and dates that weren't NA; renamed the "Pride?" variable for better readability; and looked at the total contribution for both pride and anti-LGBTQ+ campaigns over the years 2013-2022. It is notable how there was no data on contributions towards anti-LGBTQ+ campaigns in 2016.

```{r question-2-data-processing}
company_contribution_all_q2 <- contribution_data_all_states %>%
  filter(!is.na(Company), !is.na(Date)) %>%
  mutate(year = year(ymd(Date)),
         pride_or_not = case_when(`Pride?` == TRUE ~ "Pride",
                                TRUE ~ "Anti-LGBTQ+")) %>%
  group_by(pride_or_not, year) %>%
  summarise(total_contribution = sum(Amount), .groups = 'drop')

company_contribution_all_q2
```

# Analysis plan

## Plan for Answering Question 1

For the first question, we plan to visualize the answer by creating a bar chart of the dollar amount of contributions to anti-LGBTQ+ campaigns made by the top 10 (tentative) companies in USD. The x variable would be the contribution amount, the y variable would be the company names, we plan to color the bar of the company based on whether or not they have donated to pride, and we plan to add a pattern (ex. stripes with ggpattern) on the bar to indicate whether a company has an HRC pledge in place. Variables used: Company (company name), amount_contributed (contribution to anti-LGBTQ+ in USD), hrc_pledge (indicating existence of an HRC pledge, and pride_or_not (whether or not company has donated to pride).

Source: https://stackoverflow.com/questions/62393159/how-can-i-add-hatches-stripes-or-another-pattern-or-texture-to-a-barplot-in-ggp

## Plan for Answering Question 2

For the second question, we plan to create a line plot that displays the dollar amount of contributions made to both anti-LGBTQ+ campaigns and Pride campaigns by companies who contribute to both over the years of 2013 to 2022. To create the "year" variable, we used the "Date" variable originally in the dataset and extracted the year using the lubridate package. For our plot, the x variable would be the year, the y variable would be the contribution amount, and we would have two lines, one representing the contribution for anti-LGBTQ+ campaigns and the other representing the contribution for pride campaigns; we could either make the linetypes of the two lines different or have different colors for easy readability. Variables used: Contribution to anti-LGBTQ+ campaigns (USD), Contribution to Pride campaigns (USD), and Year.


```{r}
contribution_data_all_states %>%
  filter(!is.na(Company)) %>%
  mutate(year = year(Date)) %>% 
  filter(year == 2021) %>% 
  group_by(Company, `Pride?`) %>% 
  mutate(amount_cont = sum(Amount)) %>%
  relocate(amount_cont) %>% 
  select(Company, amount_cont, `Pride?`) %>%
  distinct() %>%
  arrange(desc(amount_cont)) %>% 
  head(20) %>%
  ggplot(aes(x = Company, y = amount_cont, fill = `Pride?`)) +
  geom_histogram(stat = "identity") +
  coord_flip()
```

