---
title: "Project 1: Co-Medians"
subtitle: "Project 1 Proposal"
authors: "Ellen Zhang, Amber Potter, Helen Chen, Morgan Bernstein"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
library(dplyr)
library(tidytuesdayR)
```

# Dataset

```{r}
#| label: load-data
#| message: false

pride_aggregates <- pride_aggregates
fortune_aggregates <- fortune_aggregates
static_list <- static_list
pride_sponsors <- pride_sponsors
corp_by_politician <- corp_by_politician
```

We have multiple datasets that we are planning to work with for this project from. The first one is pride_aggregates, which has 31 observations of 4 variables. It has information on companies that are pride sponsors but also contribute to anti-LGBTQ+ corporations. The variables given are the company name, total amount contributed to anti-LGBTQ+ corporations, number of politicians contributed to, and number of states where they made contributions. The second dataset, fortune_aggregates, has 117 observations of the same 4 variables but this time looking at Fortune 500 pride sponsors who donate to anti-LGBTQ+ campaigns. The third dataset is static_list, containing information on 126 companies and whether they donated to pride, HRC business pledge, dollar amount contributed across states, number of politicians contributed to, and number of states where they made contributions.

Make sure to load the data and use inline code for some of this information.

# Questions

The two questions you want to answer.

# Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
