---
title: "Project 1: Co-Medians"
subtitle: "Project 1 Proposal"
authors: "Ellen Zhang, Amber Potter, Helen Chen, Morgan Bernstein"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
library(dplyr)
library(tidytuesdayR)
library(lubridate)
```

# Dataset

```{r}
#| label: load-data
#| message: false

tuesdata <- tidytuesdayR::tt_load('2022-06-07')

pride_aggregates <- tuesdata$pride_aggregates
fortune_aggregates <- tuesdata$fortune_aggregates
static_list <- tuesdata$static_list
pride_sponsors <- tuesdata$pride_sponsors
corp_by_politician <- tuesdata$corp_by_politician
contribution_data_all_states <- tuesdata$contribution_data_all_states

company_contribution_all_q2 <- contribution_data_all_states %>%
  filter(!is.na(Company), !is.na(Date)) %>%
  mutate(year = year(ymd(Date)),
         pride_or_not = case_when(`Pride?` == TRUE ~ "Pride",
                                TRUE ~ "Anti-LGBTQ+")) %>%
  group_by( pride_or_not, year) %>%
  summarize(total_contribution = sum(Amount)) %>%

unique(contribution_data_all_states$`Pride?`)


#unique(contribution_data_all_states$Date == NA)

#glimpse(pride_aggregates)
#glimpse(fortune_aggregates)
#glimpse(static_list)
#glimpse(pride_sponsors)
#glimpse(corp_by_politician)
#glimpse(contribution_data_all_states)
```

```{r}
#Source: Politico (https://www.politico.com/2020-election/results/president/)
#by how they voted in 2020 presidential election

blue_states <- c("NJ", "CO", "AZ", "GA", "WA", "OR", "CA", "NV", "NM", "MN", 
                 "HI", "MI", "IL", "WI", "VA", "MD", "PA", "NY", "VT", "NH", 
                 "MA", "ME", "CT", "RI", "DE", "DC")


contribution_data_all_states %>% 
  filter(!is.na(Company)) %>%
  mutate(year = year(ymd(Date))) %>%
  filter(year == 2021) %>%
  mutate(party = ifelse(State %in% blue_states, "Blue", "Red")) %>%
  group_by(State) %>%
  summarise(sum_of_contributions = sum(Amount))


unique(contribution_data_all_states$State)
```

The datasets we are using are from TidyTuesday and come from the website Data For Progress. The first dataset is pride_aggregates, which has 31 observations of 4 variables. It has information on companies that are pride sponsors but also contribute to anti-LGBTQ+ corporations. The variables given are the company name, total amount contributed to anti-LGBTQ+ corporations, number of politicians contributed to, and number of states where they made contributions. The second dataset, fortune_aggregates, has 117 observations of the same 4 variables but this time looking at Fortune 500 pride sponsors who donate to anti-LGBTQ+ campaigns. The third dataset is static_list, containing information on 126 companies and whether they donated to pride, HRC business pledge, dollar amount contributed across states, number of politicians contributed to, and number of states where they made contributions. Lastly, the largest dataset is contribution_data_all_states with 13210 observations of companies along with if they were a pride and sponsor match, if they were a pride event sponsor, HRC business pledge, donor name, politician donated to, state donated to, amount in USD of donation, date, citation, donor type, comments, and an archive.


#color by HRC business pledge, x axis is contributions, y is the company 

```{r}
unique(static_list$Company)

static_list %>%  
  filter(!is.na(Company), !is.na(`HRC Business Pledge`)) %>%
  select(Company, `Amount Contributed Across States`, `HRC Business Pledge`) %>%
  group_by(Company) %>%
  summarise(total_amount = sum(`Amount Contributed Across States`))
  
```

```{r making_year_variable}
contribution_data_all_states |> 
  mutate(year = year(ymd(Date))) |> 
  distinct(year) #years go 2013-2021
```

# Questions

The two questions you want to answer.

# Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
