---
title: "Project 1: Co-Medians"
subtitle: "Project 1 Proposal"
authors: "Ellen Zhang, Amber Potter, Helen Chen, Morgan Bernstein"
format: html
editor: visual
---

```{r load-packages}
#| label: load-pkgs
#| message: false
#| warning: false

library(tidyverse)
library(dplyr)
library(tidytuesdayR)
library(lubridate)
library(readr)
```

# Dataset

```{r load-data}
#| label: load-data
#| message: false

# read data 
static_list <-readRDS("data/static_list.RDS")
contribution_data_all_states <-readRDS("data/contribution_data_all_states.RDS")
```

The datasets we are using are from TidyTuesday (https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-06-07) and come from the website Data For Progress (https://www.dataforprogress.org/accountable-allies). For context, Data For Progress is a think tank and polling firm that provides data for progressive campaigns, causes, and candidates. Their work has been used by US representatives, senators, and local leaders, and it has been cited by the White House. The datasets given were `static_list`, `contribution_data_all_states`, `pride_aggregates`, `fortune_aggregates`, `pride_sponsors`, `donors`, and `corp_by_politician`. After looking into the datasets further, we realized that there was overlap of information in `static_list`, `fortune_aggregates`, and `pride_aggregates`, where `static_list` captured the most complete information from the other two combined, so we decided to use that dataset. Our decision to also use the `contribution_data_all_states` dataset was made for similar reasons. In many ways it is a combination of the others in terms of information and variables, but we appreciated that it included that information by individual contribution rather than by company or politician.

The first dataset we are using is `static_list`, containing information on `r nrow(static_list)` companies (each row corresponds to a company) and whether they donated to pride, if they have an HRC business pledge, the dollar amount contributed across states (USD), the number of politicians contributed to, and the number of states where they made contributions. It has `r ncol(static_list)` columns/variables. A few of the variables we are using from this dataset for our analysis include `Pride?`, `HRC Business Pledge`, and `Amount Contributed Across States`. The second dataset is `contribution_data_all_states` with `r nrow(contribution_data_all_states)` observations with each row denoting a single donation/contribution, along with information on the company it was made by, whether or not they were a pride and sponsor match, if they were a pride event sponsor, HRC business pledge, donor name, politician donated to, state donated to, amount of donation (USD), date, citation, donor type, comments, and an archive. It has a total of `r ncol(contribution_data_all_states)` columns/variables. Some of the variables we are using to answer our research questions from this dataset are `Politician`, `Date` (which we extracted the column `Year` from), and `Amount`.

# Reason for Choosing this Dataset

Overall, we thought that these datasets would be extremely interesting both to visualize and to learn about. We are all advocates for LGBTQ+ rights, and we were very intrigued by the prevalence of corporate pride sponsors who also donate to anti-LGBTQ+ campaigns. This article (https://19thnews.org/2022/06/companies-anti-lgbtq-bills-lawmakers/) outlines the fact that many companies have publicly opposed anti-LGBTQ+ bills and have pro-LGBTQ campaigns, however, some of these companies simultaneously fund lawmakers who back anti-LGBTQ+ bills. In particular, from 2021-2022, the newspaper "Popular Information" found that 25 companies who were publicly in support of the pride movement have also donated \$13M to politicians who back anti-LGBTQ+ bills. We believe this is sneaky and dishonest behavior by corporations that deserves attention in order to help others become more aware of these hypocritical practices.

# Questions and Analysis Plan

## Question 1

The first question we want to answer using this dataset is: What is the relationship between a company's status as a pride sponsor, whether they have an HRC pledge, and their contributions to anti-LGBTQ+ campaigns? The Human Rights Campaign (HRC) business pledge is a commitment for companies to show their clear opposition to harmful legislation aimed at restricting the access of LGBTQ people in society and the workplace. We would expect that companies who support the LGBTQ+ community would not also make anti-LGBTQ+ donations, but this is not the case. Even more so, even companies with HRC pledges make these anti-LGBTQ+ donations. Therefore, we wanted to compare company contribution amounts to uncover patterns associated with these categorizations.

Source: https://www.hrc.org/resources/business-statement-on-anti-lgbtq-state-legislation

## Plan for Answering Question 1

For our first question, we plan to make a bar chart visualization of the top 10 or 20 companies who contributed the most to anti-LGBTQ+ campaigns (in USD). From there, we will color the bars based on whether or not the company is a Pride sponsor, and add a pattern (ex. stripes with ggpattern) on the bar to indicate whether a company has an HRC pledge in place. We are interested in seeing which companies are donating the most money to anti-LGBTQ+ campaigns, as well as how many of the top corporate contributers are blatantly anti-LGBTQ+ vs. those that are hypocritical Pride sponsors. For the second visualization, we want to investigate the distribution of anti-LGBTQ+ contribution amounts by company within the three categories of company: not a pride sponsor company, pride sponsor company without HRC pledge, pride sponsor company with HRC pledge. To accomplish this, we plan to make a density plot of the amount contributed to anti-LGBTQ+ campaigns, filled by which category of company the donation was made by. This visualization will provide insight into the distribution of contribution amounts that companies contribute to anti-LGBTQ+ campaigns across these three categories of company.

Source: https://stackoverflow.com/questions/62393159/how-can-i-add-hatches-stripes-or-another-pattern-or-texture-to-a-barplot-in-ggp

## Question 2

Our second question asks: how have the donations to politicians who support anti-LGBTQ+ campaigns changed over time? We are very curious as to how the total amount of money contributed to anti-LGBTQ+ politicians has changed from 2013-2022, as there has been a wide discussion about human rights issues especially in the past few years and we are hopeful to see a decline in the amount contributed. Speaking of the dialogue on these topics, Greg Abbott has been an outspoken Texas Republican who has received the highest number (USD) of individual donations. Because of this, we are particularly interested in the proportion of donations to anti-LGBTQ+ campaigns over the years 2019-2022 (years with data on contributions from Greg Abbott) that were made specifically to Greg Abbott. If this proportion is found to be insignificant compared with the entire contribution amounts, we will focus solely on Texas donations.

```{r}
#| label: years_greg_abbot

contribution_data_all_states %>%
  filter(Politician == "Greg Abbott") %>% 
  mutate(year = year(ymd(Date))) %>%
  select(year) %>%
  distinct(year)
```

## Plan for Answering Question 2

First, we plan to create a simple line plot showing how contributions to anti-LGBTQ+ politician campaigns (in USD amounts) has changed over time from 2013 to 2022. We feel as though this is one of the first and most basic questions that someone would have regarding data on contributions to anti-LGBTQ+ politician campaigns over multiple years. Diving further, we wanted to look individually at Greg Abbott, the receiver of the highest USD amount of contributions over all years present in our data (even though he only received contributions in the 4 years from 2019-2022). Overall, we would hypothesize that the overall trend of anti-LGBTQ+ political campaign donations would decrease, however due to the large media presence of Greg Abbott and his strong rhetoric involving this topic, we believe he might prove a contrast to the overall trend. We plan to zoom into the four years of 2019-2022 specifically and visualize the proportion of total contributions to anti-LGBTQ+ political campaigns that were made specifically to him using a stacked bar chart with argument `position` = "fill".
